Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-17 background_C.tex
+

\end{tikzpicture}
          \caption{Accessing to the Page Referenced by L1 Entry}
        \label{fig:enter-label}
    \end{subfigure}
    \vfill
\begin{minipage}{.5\textwidth}
%    \begin{figure}
\begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\footnotesize]
/*
@param entry: entry whose frame address is used 
              to locate the subsequent entry in the page-tree
@return next: return a virtual pointer to the descendant tree
*/
pte_t *pte_get_next_table(pte_t *entry) {
  pte_t *next;
  /*Reading page table entry:
    pointed by a virtual memory address*/
  /* Jump if the present bit is not zero:
     no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
  if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
    /*Entry value updates:*/
    entry->pfn = nextpaddr;
    entry->present = 1; (*\label{line:install_new_entryC}*)
    /*Entry is initialized,
      so we satisfy the condition to access children list */
  } (*\label{line:end_of_allocation_pathCB}*)
  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceSB}*)*/
  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vCB}*);
  next = (pte_t *) next_virt_addr;(*\label{line:finalpieceEB}*)
  return next;
}
\end{lstlisting}
\end{minipage}
\begin{minipage}{.4\textwidth}
\begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\footnotesize]
pte_t *walkpgdir(pte_t *l4, void *va){ 
  pte_t *l4_entry = &l4[L4Offset(va)];
  pte_t *l3 = pte_nxt_table(l4_entry);
  pte_t *l3_entry = &l3[L3Offset(va)];
  pte_t *l2 = pte_nxt_table(l3_entry);
  pte_t *l2_entry = &l2[L2Offset(va)];  
  pte_t *l1 = pte_nxt_table(l2_entry);
  pte_t *l1_entry = &l1[L1Offset(va)];  
  return l1_entry
}
\end{lstlisting}
%          \caption{Walking page-table tree to locate L1 entry C-source code.}
%        \label{fig:enter-label}
 %   \end{figure}
\end{minipage}
%\end{framed}
    \vspace{-2.5em}
    \caption{x86-64 page table lookups.}
    \label{fig:pagetables}
    \vspace{-1em}
\end{figure}

>>>>

\end{tikzpicture}
          \caption{Accessing to the Page Referenced by L1 Entry}
        \label{fig:enter-label}
    \end{subfigure}
    \vfill
\begin{minipage}{.5\textwidth}
%    \begin{figure}
\begin{lstlisting}[mathescape,escapeinside={(*}{*)},basicstyle=\footnotesize]
/*
@param entry: entry whose frame address is used 
              to locate the subsequent entry in the page-tree
@return next: return a virtual pointer to the descendant tree
*/
pte_t *pte_get_next_table(pte_t *entry) {
  pte_t *next;
  /*Reading page table entry:
    pointed by a virtual memory address*/
  /* Jump if the present bit is not zero:
     no need to allocate next level (* \label{line:check_entry_presentC} *)(* \label{line:mask_presentC} *)  (*\label{line:after_concluding_qfrac1C}*)*/
  if (!entry->present(*\label{line:read_entry_contentsC}*) (*\label{line:conditional_childrenC}*)) {
    /*Allocation path starts here (*\label{line:pass_addrof_nextC}*)(*\label{line:alloc_path_startC}*)*/
    pte_initialize(entry);(* \label{line:call_to_pte_initializeC} *)
    /*Entry value updates:*/
    entry->pfn = nextpaddr;
    entry->present = 1; (*\label{line:install_new_entryC}*)
    /*Entry is initialized,
      so we satisfy the condition to access children list */
  } (*\label{line:end_of_allocation_pathCB}*)
  /*Physical to virtual address mapping code segment starts here (*\label{line:finalpieceSB}*)*/
  uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn)
  uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr); (*\label{line:p2vCB}*);
  next = (pte_t *) next_virt_addr;(*\label{line:finalpieceEB}*)
  return next;
}
\end{lstlisting}
\end{minipage}
\begin{minipage}{.4\textwidth}
\begin{lstlisting}[mathescape,escapeinside={(*}{*)}, basicstyle=\footnotesize]
pte_t *walkpgdir(pte_t *root, void *va){ 
  pte_t *l4_entry = &root[L4Offset(va)];
  pte_t *l3 = pte_next_table(l4_entry);
  pte_t *l3_entry = &l3[L3Offset(va)];
  pte_t *l2 = pte_next_table(l3_entry);
  pte_t *l2_entry = &l2[L2Offset(va)];  
  pte_t *l1 = pte_next_table(l2_entry);
  pte_t *l1_entry = &l1[L1Offset(va)];  
  return l1_entry
}
\end{lstlisting}
%          \caption{Walking page-table tree to locate L1 entry C-source code.}
%        \label{fig:enter-label}
 %   \end{figure}
\end{minipage}
%\end{framed}
    \vspace{-2.5em}
    \caption{x86-64 page table lookups.}
    \label{fig:pagetables}
    \vspace{-1em}
\end{figure}
